On The Impact of Atoms of Confusion in JavaScript Code

### 2021-09-27 4_Result.tex
Updates from Overleaf
final review

Our first study investigates the impact of atoms of confusion while developers try to understand JavaScript code.
We estimate this impact considering two perspectives: \emph{misunderstanding rate} (number of wrong answers)
and \emph{cognitive effort} (time necessary to provide a correct answer). 
We received full answers to our survey from 140 participants (a total of 70 replicas).
All participants had taken at least some university course or hold a bachelor
degree or equivalent, meaning the participants have had some level of formal education in programming. Twenty four participants hold either a master's degree (21 participants) or a doctorate degree (three participants). 
Considering the programming experience of our respondents, 19\% have more than ten years of programming experience, 37\% have between four and ten years of experience, 37\% have between one and four years of experience, and 7\% have less than one year of programming experience.  
Accordingly, we characterize the effect of atoms of confusion in JavaScript
code taking into account the perceptions of both novice and experienced developers.

@@

As mentioned in Section~\ref{sec:meth:survey}, each participant in this study evaluated ten
code snippets, from which five were in their confusing versions, whilst the other five contained clean versions of the code snippets (i.e., without the confusing constructs and idioms). The participants should provide the expected outcomes of the code snippets. As discussed, we collected information about \emph{correctness} (whether the participant correctly predicted the program's output) and \emph{cognitive workload} (the time taken to answer the question). 
Table~\ref{tab:difference-correctness} and Figure~\ref{fig:boxplotcorrectness} summarize the results of the survey w.r.t. correctness.

@@

\begin{table}[htbp]
\caption{Total number of correct answers}
\label{tab:difference-correctness}
\centering{
\begin{scriptsize}
\begin{tabular}{lccc} \toprule
  Atom & Confusing Code & Clean Code & $\Delta$(\%)  \\ \midrule
  Comma Operator                &  28 &  65 & +132   \\ 
  Automatic Semicolon Insertion &  32 &  68 & +112   \\ 
  Post-Increment                &  48 &  64 & +33    \\ 
  Omitted Curly Braces          &  47 &  58 &  +23   \\
  Assignment as Value           &  56 &  68 &  +21   \\ 
  Implicit Predicate            &  58 &  68 &  +17   \\ 
  Logic as Control Flow         &  41 &  48 &  +17   \\ 
  Ternary Operator              &  60 &  66 &  +10   \\ 
  Pre-Increment                 &  50 &  53 &  +6    \\ 
  Arithmetic as Logic           &  64 &  63 &  -2    \\ \bottomrule

@@

\begin{figure}[htb!]
\noindent
 \centering
 \includegraphics[width=\columnwidth]{images/wrong-answers-plot-1.pdf}
 \caption{Number of wrong answers of each subject.}
 \label{fig:boxplotcorrectness}
 \end{figure}

@@

We measure the effect size of the clean version of the code snippets into the answers' correctness using the \emph{Odds Ratio} (OR). The results are reported in the ``Odds Ratio'' column of Table~\ref{tab:hypothesis-testing}. In the table, we also report the confidence interval (CI) for the OR in the ``CI'' column. Although many of the intervals are wide, for six of the atom candidates the lower bound of the confidence interval is greater than or equal to 1. This indicates that, at a 95\% confidence level, a developer is likely to commit less mistakes when using the clean versions.
%We found a negligible effect size for the atom candidates Arithmetic as Logic, Logic as Control Flow, and Pre-Increment. Nonetheless, for the remaining candidates, the effect size is significant. 
For instance, when comparing clean and confusing code snippets pertaining to the Comma Operator atom candidate, we observed an \emph{Odds Ratio} of \num{19.02}. This means that the odds of a correct answer are \num{19.02} times higher when interpreting the clean version---in comparison with the corresponding confusing version of the code snippet. Furthermore, with a 95\% confidence level, the odds  ratio is between 6.6 and 68.22, i.e., although the error margin is wide, it is strongly in favor of the clean version. If there is a significant likelihood of a developer committing a mistake when analyzing a clean version, the lower bound of the CI should be a number between 0 and 1 (since the OR is a ratio). This is the case for the four atom candidates in the lower part of the table. In particular, in the last row (Arithmetic as Logic), the OR between 0 and 1 indicates that a study subject was actually more likely to misunderstand the clean version.

@@

\subsection{Cognitive Workload Analysis}

\subsubsection*{Exploratory Data Analysis}
Table \ref{tab:difference-time-taken} shows the average time necessary for the participants to give a correct answer about the expected outcomes of a code snippet, considering both confusing and clean versions. Seven atom candidates require an increasing cognitive workload. For these seven atom candidates, developers take at least 5.90\% less time on average to find a correct answer when considering the clean version of a code snippet. In the extreme case (atom candidate Comma Operator), the participants take 76.23\% less
time on average to find the correct answer for the clean version of the code snippets. 
Not all atom candidates, though, require less time for the participants to predict the answer. In fact, for the Arithmetic as Logic and the Pre-Increment atoms, the time to give a correct answer increased  
29.06\% and 38.19\% for the clean versions.

\begin{table}[htbp]
\caption{Time in seconds to submit a correct answer}
\label{tab:difference-time-taken}
\centering{
\begin{scriptsize}
\begin{tabular}{lccc} \toprule
Atom & Confusing Code & Clean Code & $\Delta$(\%) \\ \midrule
Comma Operator & 87.67 & 20.84 & -76.23 \\ 
Automatic Semicolon Insertion & 46.08 & 22.04 & -52.17 \\ 
Post-Increment & 28.70 & 25.67 & -10.56 \\ 
Omitted Curly Braces & 48.85 & 30.00 & -38.58 \\ 
Assignment as Value & 52.47 & 48.95 & -6.71 \\ 
Implicit Predicate & 36.24 & 24.01 & -33.75 \\ 
Logic as Control Flow & 108.94 & 51.07 & -53.12 \\ 
Ternary Operator & 41.80 & 39.34 & -5.90 \\ 
Pre-Increment & 30.71 & 42.45 & 38.19 \\  
Arithmetic as Logic & 28.82 & 37.20 & 29.06 \\ \bottomrule
\end{tabular}
\end{scriptsize}
}
\end{table}

 % latex table generated in R 4.0.4 by xtable 1.8-4 package
 % Thu Apr 22 07:58:21 2021
 \begin{table*}[ht]
\caption{Hypotheses Testing (misunderstanding and cognitive workload)}
\label{tab:hypothesis-testing}

 \centering
 \begin{tabular}{lccc|cc}
   \toprule
Atom  & Chi-square test & Odds Ratio          & CI & Mann-Whitey U test & Cliff's Delta \\ 
     &         & (misunderstanding)  &                    & (workload)             & \\ \midrule
Comma Operator & 1.1727e-10 & 19.02 & (6.60, 68.22) & 5.8249e-08 & -0.53 \\ 
Automatic Semicolon Insertion & 5.8352e-11 & 39.33 & (9.21, 356.62) & 0.09802 & -0.16 \\ 
Post-Increment & 0.0015279 & 4.83 & (1.73, 15.74) & 0.12006 & 0.15 \\ 
Omitted Curly Braces & 0.050962 & 2.35 & (1.00, 5.77) & 0.92529 & -0.01 \\ 
Assignment as Value & 0.0034775 & 8.39 & (1.81, 79.12) & 0.82681 & 0.02 \\ 
Implicit Predicate & 0.01123 & 6.95 & (1.46, 66.56) & 0.0029039 & -0.29 \\ 
Logic as Control Flow & 0.292 & 1.54 & (0.73, 3.28) & 7.5862e-05 & -0.39 \\ 
Ternary Operator & 0.15896 & 2.73 & (0.74, 12.57) & 0.24073 & -0.12 \\ 
Pre-Increment & 0.70147 & 1.25 & (0.55, 2.85) & 0.0062981 & 0.27 \\ 
Arithmetic as Logic & 1 & 0.84 & (0.22, 3.12) & 0.01723 & 0.23 \\ 
    \bottomrule
 \end{tabular}
 \end{table*}

\subsubsection*{Hypothesis Testing}
We use the non-parametric \emph{Mann-Whitney U test}, also known as Wilcoxon rank-sum test, to
investigate the null hypothesis that developers 
spend the same amount of time to correctly
predict the outcome of a code snippet, regardless
of evaluating the confusing or  clean version
of the code. The results of Table~\ref{tab:difference-time-taken}
suggest that we should refute the null hypotheses for
six atom candidates (Comma Operator, Automatic Semicolon Insertion,
Implicit Predicate, Logic
as Control Flow, Pre-Increment, and Arithmetic
as Logic). For the first four, the analysis suggests that the participants
need more time to predict the outcome of
the code snippet in the confusing code version. Interestingly, for the atom candidates Arithmetic as Logic and Pre-Increment, the participants take less time
to predict the outcome of the code snippets in
the confusing version.
We also computed the effect size using the Cliff's Delta method. We found a large effect for the Comma Operator atom candidate; and a moderate effect for the atom candidates
Logic as Control Flow, Implicit Predicate, Arithmetic
as Logic, and Pre-Increment.

\begin{mh}
  The results suggest that developers spend
  significantly more time to predict the
  correct answer in the presence of the
  atom candidates Comma Operator, Automatic Semicolon Insertion, Logic as
  Control Flow, and Implicit Predicate. Conversely,
  the code snippets with the
  atom candidates Arithmetic as Logic and
  Pre-Increment demand less time to
  predict the correct answers. 
\end{mh}

@@

\section{Results of the Interview Study}
\label{sec:interview-results}

@@

In this section, we present the results of the 
interviews with the 15 practitioners who agreed to participate
in the study. 
% A total of 15 practitioners took part of the interviews.
We collected information regarding programming
experience, familiarity with JavaScript, and their opinion
about the \na atom candidates we explored in our research.
We first present them pairs of clean and confusing code snippets 
and then ask them to discuss their preferences towards one version.
% ask the participants to discuss their preferences (i.e., code with or without atom).

@@

\subsubsection*{The Participants' perceptions of the atom candidates}
Supporting the results of the survey, Table~\ref{tab:interview-results1}
 shows that for eight out of the
10 scenarios surveyed, the majority of the respondents prefer the version of the
code without the atom of confusion.
In no case the \emph{neutral}
ratio was higher than the option for the version without atoms of confusion.
%% An example entry is contained in Appendix~\ref{}. 
%% For a full listing of the code snippets, visit the paper repository. 
Only for the Ternary Operator atom candidate the participants
prefer the version of the code with the atom of confusion,
instead of the version without the atom of confusion.
We show the code snippets for the Ternary Operator in Figure~\ref{code:ternary}.
Even for this atom candidate, some participants who preferred 
the \emph{left-hand-side} version (with the atom of confusion) 
still believed that the \emph{right-hand-side} version (without the atom of
confusion) was more readable.  
The following quotes were extracted from the transcripts with
% three interviewees:
two interviewees:

@@

\noindent\begin{minipage}{.45\textwidth}
\begin{lstlisting}[language=JavaScript, caption=\emph{Left-hand side} (using the \emph{Ternary Operator} atom).]
let config = {size: 3, isActive: false};
const_config = config.isActive === true 
             ? config 
             : {size: 10};
console.log(_config.size);
\end{lstlisting}
\end{minipage}\hfill
\begin{minipage}{.45\textwidth}
\begin{lstlisting}[language=JavaScript, caption=\emph{Right-hand side} (without the atom).]
let config = {size: 3, isActive: false}
let _config;
if(config.isActive === true) {
  _config = config;
}
else{
   _config = {size: 10};
}
console.log(_config.size);
\end{lstlisting}
\end{minipage}
\caption{Example of pair of code snippets used in the interview. This case explores the use of the \emph{Ternary Operator atom of confusion}}
\label{code:ternary}
\end{figure*}

@@

\begin{mq}
\emph{``When I am programming, I write code with the ternary operator, [...], but, to be honest, I still think that the 
[the code using the \lhs version] 
is easier to understand''}.
\end{mq}

@@

The Pre-Increment atom candidate also caused such
divide in one of the interviewees, as they
regarded the \emph{\rhs version} (without atom of confusion)
as simpler to understand, but would still opt to write
code with the atom. In contrast, %to such opinion,
one of the participants found the 
version with the atom of confusion more elegant,
but recognized it was less readable,
and was willing to sacrifice elegance for readability.

@@

When analyzing the Logic as Control Flow atom candidate, one of
the interviewees gave an example of his own experience
that might motivate one to avoid writing code using this construct:

\begin{mq}
  \emph{``This one is interesting, because I have written
  code that looks like the [\lhs version (with the atom of confusion)],
  and my colleagues complained that it was difficult to understand.
  Nowadays I prefer to write code using the [\rhs version]"}.
\end{mq}

Two of the atom candidates had unanimous preference for the
versions that did not contain the atoms: Comma Operator and
Omitted Curly Braces. Regarding the first one, we could
often notice during the interviews that
the \emph{\lhs} (with the atom of confusion) 
caused significant
confusion among the participants. One remark about
the \emph{Comma Operator} atom of confusion is listed below:

@@

As for the atom candidate Omitted Curly Braces, one of the interviewees mentioned that, although they understand why one would opt not to use braces for simple if-then-else statements, he still advised against it, on grounds that:

@@

This is an interesting perspective because such a perhaps simple decision might hinder novice developers in maintaining code; one cannot assume the programmers' experience of who will maintain a code in the long run.

\begin{table}[!htb]
    \centering
    \caption{Summary of participants' preferences for code snippets \emph{with} (LHS) and
      \emph{without} atoms of confusion (RHS).
      Participants were also allowed to choose NEUTRAL when they thought both sides were equally readable.}
    \label{tab:interview-results1}
    \begin{tabular}{lrrr}\toprule
      & \multicolumn{3}{c}{\textsc{Preference (\%)}} \\
      \cmidrule(lr){2-4}
         Atom           & \multicolumn{1}{c}{\textsc{LHS}}
                                      &  \multicolumn{1}{c}{\textsc{RHS}}
                                               & \multicolumn{1}{c}{NEUTRAL} \\ \midrule
         Comma Operator                  & 0  & 100    & 0     \\
         Automatic Semicolon Insertion  & 0  & 80     & 0     \\
         Post-Increment                  & 20 & 73.33  & 6.67  \\
         Omitted Curly Braces            & 0  & 100    & 0     \\
         Assignment as Value             & 20 & 60     & 20    \\
         Implicit Predicate              & 20 & 73.33  & 6.67  \\
         Logic as Control Flow           & 20 & 60     & 20    \\
         Ternary Operator                & 60 & 26.67  & 13.3  \\
         Pre-Increment                   & 40 & 46.67  & 13.33 \\ 
         Arithmetic as Logic             & 0  & 93.33  & 6.67  \\ \midrule
         \textsc{overall}                & 18 & 71.33  & 8.64  \\
         \bottomrule
    \end{tabular}
\end{table}

@@

\begin{figure}
\begin{small}
\begin{lstlisting}[language=JavaScript,caption=Example of arrow function.,label=arr1]
let inc = (x) => x + 1;
\end{lstlisting}
\begin{lstlisting}[language=JavaScript,caption=Alternative version without arrow function.,label=arr2]
let inc = function(x) {
  return x + 1
}
\end{lstlisting}
\end{small}
\end{figure}

@@

We mined \minedprojects open source JavaScript repositories to understand how
often atoms of confusion arise in real software. 
Similarly to previous studies~\cite{DBLP:conf/msr/GopsteinZFC18,DBLP:journals/ese/MedeirosLAAKRG19}, which investigate
the prevalence of atoms of confusion in open source \clang and \cpplang projects, we found that atoms of confusion
frequently arise in JavaScript open source systems. For instance, the six most frequently found atoms
occur in at least 80\% of the projects. Considering the extremes, atoms Ternary Operator and Implicit Predicate were found in all repositories we mined,
while atom Comma Operator occurred in only 15\% of them (as seen in Table~\ref{tab:occurrences-summary}).

@@

% latex table generated in R 4.0.3 by xtable 1.8-4 package
% Tue Apr 13 09:34:35 2021
\begin{table}[ht]
\centering
\caption{Summary of atoms occurrences on our dataset}
\setlength\tabcolsep{2pt} % default value: 6pt
\label{tab:occurrences-summary}
\begin{tabular}{lrcc}
  \toprule
Atom & Projects & Occurr./KLOC & Frequency \\ 
  \midrule
Implicit Predicate & 100\% & 19.89 &  highly used \\ 
  Ternary Operator & 100\% & 10.16 &  highly used \\ 
  Omitted Curly Braces & 91.67\% & 6.61 &  highly used \\ 
  Post Increment & 90.28\% & 5.62 &  highly used \\ 
  Pre Increment & 83.33\% & 1.04 & commonly used \\ 
  Assignment as Value & 81.94\% & 1.02 & commonly used \\ 
  Logic as Control Flow & 56.94\% & 0.95 & commonly used \\ 
  Automatic Semicolon Insertion & 33.33\% & 0.13 & commonly used \\ 
  Arithmetic as Logic & 23.61\% & 0.02 & little used \\ 
  Comma Operator & 15.28\% & 0.03 & little used \\ 
   \bottomrule
\end{tabular}
\end{table}

Considering all JavaScript projects in our dataset, we found 
a total of \num{364873} atoms of confusion, though four atoms are responsible
for 92.97\% of this total: Implicit Predicate, Ternary Operator, Omitted Curly Braces, and the Post Increment. The remaining atoms
account for \num{25621} (7.03\%) of the total number of occurrences.
Moreover, the Comma Operator and Arithmetic as Logic atoms are the ones that arise less
frequently (0.001\% in total with 232 and 165 occurrences, respectively).

>>>>

Our first study investigates the impact of atoms of confusion while developers try to understand JavaScript code.
We estimate this impact considering two perspectives: \emph{misunderstanding rate} (number of wrong answers)
and \emph{time} (how long to provide a correct answer).
We received full answers to our survey from 140 participants (a total of 70 replicas).
All participants had taken at least some university course or hold a bachelor
degree or equivalent, meaning the participants have had some level of formal education in programming. Twenty four participants hold either a master's degree (21 participants) or a doctorate degree (three participants). 
Considering the programming experience of our respondents, 19\% have more than ten years of programming experience, 37\% have between four and ten years of experience, 37\% have between one and four years of experience, and 7\% have less than one year of programming experience.  
Accordingly, we characterize the effect of atoms of confusion in JavaScript
code taking into account the perceptions of both novice and experienced developers.

@@

As mentioned in Section~\ref{sec:meth:survey}, each participant in this study evaluated ten
code snippets, from which five were in their confusing versions, whilst the other five contained clean versions of the code snippets (i.e., without the confusing constructs and idioms). The participants should provide the expected outcomes of the code snippets. As discussed, we collected information about \emph{correctness} (whether the participant correctly predicted the program's output) and \emph{time}. 
Table~\ref{tab:difference-correctness} and Figure~\ref{fig:boxplotcorrectness} summarize the results of the survey w.r.t. correctness.

@@

\begin{table}[htbp]
\caption{Summary of the correctness analysis}
\label{tab:difference-correctness}
\centering{
\begin{scriptsize}
\begin{tabular}{lccc} \toprule
  Atom & Confusing Code & Clean Code & $\Delta$(\%)  \\ \midrule
  Comma Operator                &  28 &  65 & +132   \\ 
  Automatic Semicolon Insertion &  32 &  68 & +112   \\ 
  Post-Increment                &  48 &  64 & +33    \\ 
  Omitted Curly Braces          &  47 &  58 &  +23   \\
  Assignment as Value           &  56 &  68 &  +21   \\ 
  Implicit Predicate            &  58 &  68 &  +17   \\ 
  Logic as Control Flow         &  41 &  48 &  +17   \\ 
  Ternary Operator              &  60 &  66 &  +10   \\ 
  Pre-Increment                 &  50 &  53 &  +6    \\ 
  Arithmetic as Logic           &  64 &  63 &  -2    \\ \bottomrule

@@

\begin{figure}[b!]
\noindent
\vspace*{-1.4cm}
 \centering
 \includegraphics[width=\columnwidth]{images/wrong-answers-plot-1.pdf}
 \vspace*{-1.3cm}
 \caption{Number of wrong answers of each subject.}
 \label{fig:boxplotcorrectness}
 \end{figure}

@@

We measure the effect size of the clean version of the code snippets into the answers' correctness using the \emph{Odds Ratio} (OR). The results are reported in the ``Odds Ratio'' column of Table~\ref{tab:hypothesis-testing}. In the table, we also report the confidence interval (CI) for the OR in the ``CI'' column. Although many of the intervals are wide, for six of the atom candidates the lower bound of the confidence interval is greater than or equal to 1. This indicates that, at a 95\% confidence level, a developer is likely to commit less mistakes when using the clean versions. 
%Considering only the atoms for which there was a significant difference in correctness for confusing and clean versions and u
Using the thresholds established by Chen and colleagues~\cite{Chen:2010:HBB} (OR = 1.68 small, 3.47 medium, and 6.71 large), for four of the atoms, Comma Operator, Automatic Semicolon Insertion, Assignment as Value, and Implicit Predicate, effect size can be considered large. For Implicit Predicate, it is medium. 
%We found a negligible effect size for the atom candidates Arithmetic as Logic, Logic as Control Flow, and Pre-Increment. Nonetheless, for the remaining candidates, the effect size is significant. 
For instance, when comparing clean and confusing code snippets pertaining to the Comma Operator atom candidate, we observed an \emph{Odds Ratio} of \num{19.02}. This means that the odds of a correct answer are \num{19.02} times higher when interpreting the clean version---in comparison with the corresponding confusing version of the code snippet. Furthermore, with a 95\% confidence level, the odds  ratio is between 6.6 and 68.22, i.e., although the error margin is wide, it is strongly in favor of the clean version. If there is a significant likelihood of a developer committing a mistake when analyzing a clean version, the lower bound of the CI should be a number between 0 and 1 (since the OR is a ratio). This is the case for the four atom candidates in the lower part of the table. In particular, in the last row (Arithmetic as Logic), the OR between 0 and 1 indicates that a study subject was actually more likely to misunderstand the clean version.

@@

\subsection{Time Analysis}

\subsubsection*{Exploratory Data Analysis}
Table \ref{tab:difference-time-taken} shows the average time necessary for the participants to give a correct answer about the expected outcomes of a code snippet, considering both confusing and clean versions. We do not consider cases where the participants made mistakes. Seven atom candidates required more time from participants to predict a correct response. For these seven atom candidates, developers take at least 5.90\% less time on average to find a correct answer when considering the clean version of a code snippet. In the extreme case (atom candidate Comma Operator), the participants take 76.23\% less
time on average to find the correct answer for the clean version of the code snippets. 
Not all atom candidates, though, require less time for the participants to predict the answer. In fact, for the Arithmetic as Logic and the Pre-Increment atoms, the time to give a correct answer increased  
29.06\% and 38.19\% for the clean versions.

\begin{table}[htbp]
\caption{Time in seconds to submit a correct answer}
\centering{
\begin{scriptsize}
\begin{tabular}{lccc} \toprule
Atom & Confusing Code & Clean Code & $\Delta$(\%) \\ \midrule
Comma Operator & 87.67 & 20.84 & -76.23 \\ 
Automatic Semicolon Insertion & 46.08 & 22.04 & -52.17 \\ 
Post-Increment & 28.70 & 25.67 & -10.56 \\ 
Omitted Curly Braces & 48.85 & 30.00 & -38.58 \\ 
Assignment as Value & 52.47 & 48.95 & -6.71 \\ 
Implicit Predicate & 36.24 & 24.01 & -33.75 \\ 
Logic as Control Flow & 108.94 & 51.07 & -53.12 \\ 
Ternary Operator & 41.80 & 39.34 & -5.90 \\ 
Pre-Increment & 30.71 & 42.45 & 38.19 \\  
Arithmetic as Logic & 28.82 & 37.20 & 29.06 \\ \bottomrule
\end{tabular}
\label{tab:difference-time-taken}
\end{scriptsize}
}
\end{table}

 % latex table generated in R 4.0.4 by xtable 1.8-4 package
 % Thu Apr 22 07:58:21 2021
 \begin{table*}[ht]
\caption{Hypotheses Testing (misunderstanding and time)}

 \centering
 \begin{tabular}{lccc|cc}
   \toprule
Atom  & Chi-square test & Odds Ratio          & CI & Mann-Whitey U test & Cliff's Delta \\ 
     &         & (misunderstanding)  &                    & (workload)             & \\ \midrule
Comma Operator & 1.1727e-10 & 19.02 & (6.60, 68.22) & 5.8249e-08 & -0.53 \\ 
Automatic Semicolon Insertion & 5.8352e-11 & 39.33 & (9.21, 356.62) & 0.09802 & -0.16 \\ 
Post-Increment & 0.0015279 & 4.83 & (1.73, 15.74) & 0.12006 & 0.15 \\ 
Omitted Curly Braces & 0.050962 & 2.35 & (1.00, 5.77) & 0.92529 & -0.01 \\ 
Assignment as Value & 0.0034775 & 8.39 & (1.81, 79.12) & 0.82681 & 0.02 \\ 
Implicit Predicate & 0.01123 & 6.95 & (1.46, 66.56) & 0.0029039 & -0.29 \\ 
Logic as Control Flow & 0.292 & 1.54 & (0.73, 3.28) & 7.5862e-05 & -0.39 \\ 
Ternary Operator & 0.15896 & 2.73 & (0.74, 12.57) & 0.24073 & -0.12 \\ 
Pre-Increment & 0.70147 & 1.25 & (0.55, 2.85) & 0.0062981 & 0.27 \\ 
Arithmetic as Logic & 1 & 0.84 & (0.22, 3.12) & 0.01723 & 0.23 \\ 
    \bottomrule
 \end{tabular}
 \label{tab:hypothesis-testing}
 \end{table*}

\subsubsection*{Hypothesis Testing}
We use the non-parametric \emph{Mann-Whitney U test}, also known as Wilcoxon rank-sum test, to
investigate the null hypothesis that developers 
spend the same amount of time to correctly
predict the outcome of a code snippet, regardless
of evaluating the confusing or clean version
of the code. The results of Table~\ref{tab:difference-time-taken}
suggests that we should refute the null hypotheses for five atom candidates (Comma Operator, Implicit Predicate, Logic as Control Flow, Pre-Increment, and Arithmetic as Logic), after applying the Benjamini-Hochberg correction with a false discovery rate of 5\%. For the first three, the analysis suggests that the participants need more time to predict the outcome of the code snippet in the confusing code version. For the atom candidates Arithmetic as Logic and Pre-Increment, the participants take less time to predict the outcome of the code snippets in the confusing version. We also computed the effect size using Cliff's Delta statistic. In the table, negative effect sizes suggest that it took less time to correctly predict the output of clean versions of the snippets.  Considering the thresholds established by Romano et al.~\cite{Romano:2006:ASO} (0.147 $<$ $|$d$|$ $<$ 0.33 small, $|$d$|$ $<$ 0.474 medium, otherwise large), we found a large effect for the Comma Operator atom candidate; a medium effect size for Logic as Control Flow, and small effect sizes for Automatic Semicolon Insertion, Implicit Predicate, Arithmetic as Logic, Post-, and Pre-Increment.

\begin{mh}
  The results suggest that developers tend to spend
  more time to predict the correct answer in the presence of the atom candidates Comma Operator, Logic as Control Flow, and Implicit Predicate. Conversely, the code snippets with the atom candidates Arithmetic as Logic and Pre-Increment demand less time to predict the correct answers. 
\end{mh}

@@

\vspace*{-0.5cm}
\section{Results of the Interview Study}
\label{sec:interview-results}

@@

In this section, we present the results of the 
interviews with the 15 practitioners who agreed to participate in the study. 
% A total of 15 practitioners took part of the interviews.
We collected information regarding programming
experience, familiarity with JavaScript, and their opinion about the \na atom candidates we explored in our research. We first presented them pairs of clean and confusing code snippets and then asked them to discuss their preferences towards one version. We did not indicate in any way whether any version was assumed to be confusing or not. 
% ask the participants to discuss their preferences (i.e., code with or without atom).

@@

\begin{table}[!htb]
    \centering
    \caption{Summary of participants' preferences for code snippets \emph{with} and
      \emph{without} atom candidates. The participants were only presented the code snippets, without any indication about whether one was confusing or not. 
      Participants were also allowed to choose Neutral when they thought both sides were equally readable.}
    \begin{tabular}{lrrr}\toprule
      & \multicolumn{3}{c}{\textsc{Preference (\%)}} \\
      \cmidrule(lr){2-4}
         Atom           & \multicolumn{1}{c}{Confusing}
                                      &  \multicolumn{1}{c}{Clean}
                                               & \multicolumn{1}{c}{Neutral} \\ \midrule
         Comma Operator                  & 0  & 100    & 0     \\
         Automatic Semicolon Insertion  & 0  & 80     & 0     \\
         Post-Increment                  & 20 & 73.33  & 6.67  \\
         Omitted Curly Braces            & 0  & 100    & 0     \\
         Assignment as Value             & 20 & 60     & 20    \\
         Implicit Predicate              & 20 & 73.33  & 6.67  \\
         Logic as Control Flow           & 20 & 60     & 20    \\
         Ternary Operator                & 60 & 26.67  & 13.3  \\
         Pre-Increment                   & 40 & 46.67  & 13.33 \\ 
         Arithmetic as Logic             & 0  & 93.33  & 6.67  \\ \midrule
         \textsc{overall}                & 18 & 71.33  & 8.64  \\
         \bottomrule
    \end{tabular}
    \label{tab:interview-results1}
\end{table}

\subsubsection*{The Participants' perceptions of the atom candidates}
Supporting the results of the survey, Table~\ref{tab:interview-results1}
 shows that for eight out of the
10 scenarios surveyed, the majority of the respondents prefer the version of the
code without the atom candidate.
In no case the \emph{neutral}
ratio was higher than the option for the Clean version.
%% An example entry is contained in Appendix~\ref{}. 
%% For a full listing of the code snippets, visit the paper repository. 
Only for the Ternary Operator atom candidate the participants prefer the Confusing version, instead of the Clean one. We show the code snippets for the Ternary Operator in Figure~\ref{code:ternary}. Even for this atom candidate, some participants who preferred the left-hand side version (confusing) still believed that the right-hand side version (clean) was more readable. The following quotes were extracted from the transcripts with
% three interviewees:
two interviewees:

@@

\noindent\begin{minipage}{.45\textwidth}
\begin{lstlisting}[language=JavaScript, caption=\emph{Left-hand side} (using the \emph{Ternary Operator} atom).]
let config = {size: 3, isActive: false};
const_config = config.isActive === true 
             ? config 
             : {size: 10};
console.log(_config.size);
\end{lstlisting}
\end{minipage}\hfill
\begin{minipage}{.45\textwidth}
\begin{lstlisting}[language=JavaScript, caption=\emph{Right-hand side} (without the atom).]
let config = {size: 3, isActive: false}
let _config;
if(config.isActive === true) {
  _config = config;
}
else{
   _config = {size: 10};
}
console.log(_config.size);
\end{lstlisting}
\end{minipage}
\caption{Example of a pair of code snippets used in the interview. This case explores the use of the \emph{Ternary Operator atom of confusion}}
\label{code:ternary}
\vspace{-0.2cm}
\end{figure*}

@@

\begin{mq}
\emph{``When I am programming, I write code with the ternary operator, [...], but, to be honest, I still think that the 
[the code using the \rhs version] 
is easier to understand''}.
\end{mq}

@@

The Pre-Increment atom candidate also caused such
divide in one of the interviewees, who
regarded the clean version as simpler to understand, but would still opted to write
code with the atom. In contrast, %to such opinion,
one of the participants found the version with the atom candidate more elegant, but recognized it was less readable, and was willing to sacrifice elegance for readability.

@@

When analyzing the Logic as Control Flow atom candidate, one of the interviewees gave an example personal experience that might motivate one to avoid writing code using this construct:

\begin{mq}
  \emph{``This one is interesting, because I have written code that looks like the left-hand side [Confusing version], and my colleagues complained that it was difficult to understand. Nowadays I prefer to write code using the version on the right-hand side [Clean]''}.
\end{mq}

For two of the atom candidates, the interviewees were unanimous in their preference for the clean version: Comma Operator and Omitted Curly Braces. Regarding the first one, we could often notice during the interviews that the \emph{\lhs} (with the atom of confusion) caused significant confusion among the participants. One remark about the \emph{Comma Operator} atom of confusion is listed below:

@@

As for the atom candidate Omitted Curly Braces, one of the interviewees mentioned that, although they understand why one would opt not to use braces for simple if-then-else statements, they still advised against it, on grounds that:

@@

%This is an interesting perspective because such a simple decision might hinder novice developers in maintaining code; one cannot assume the programmers' experience of who will maintain a code in the long run.

@@

\begin{figure}
\begin{small}
\begin{lstlisting}[language=JavaScript,caption=Example of arrow function.,label=arr1]
let inc = (x) => x + 1;
\end{lstlisting}
\begin{lstlisting}[language=JavaScript,caption=Alternative version without arrow function.,label=arr2]
let inc = function(x) {
  return x + 1
}
\end{lstlisting}
\vspace*{-0.6cm}
\end{small}
\end{figure}

@@

We mined \minedprojects open source JavaScript repositories to understand how
often atoms of confusion arise in real software. 
Similarly to previous studies~\cite{DBLP:conf/msr/GopsteinZFC18,DBLP:journals/ese/MedeirosLAAKRG19}, which investigate
the prevalence of atoms of confusion in open source \clang and \cpplang projects, we found that atoms of confusion
frequently arise in JavaScript open source systems. For instance, the six most frequently found atoms
occur in at least 80\% of the projects. Considering the extremes, atom candidates Ternary Operator and Implicit Predicate were found in all repositories we mined,
while Comma Operator occurred in only 15\% of them (as seen in Table~\ref{tab:occurrences-summary}).

@@

% latex table generated in R 4.0.3 by xtable 1.8-4 package
% Tue Apr 13 09:34:35 2021
\begin{table}[ht]
\centering
\caption{Summary of atom candidate occurrences on our dataset.}
\setlength\tabcolsep{2pt} % default value: 6pt
\begin{tabular}{lrc}%c}
  \toprule
Atom & Projects & Occurr./KLOC \\%& Frequency \\ 
  \midrule
Implicit Predicate & 100\% & 19.89 \\%&  intensively used \\ 
  Ternary Operator & 100\% & 10.16 \\%&  intensively used \\ 
  Omitted Curly Braces & 91.67\% & 6.61 \\%&  intensively used \\ 
  Post Increment & 90.28\% & 5.62 \\%&  intensively used \\ 
  Pre Increment & 83.33\% & 1.04 \\%& commonly used \\ 
  Assignment as Value & 81.94\% & 1.02 \\%& commonly used \\ 
  Logic as Control Flow & 56.94\% & 0.95 \\%& commonly used \\ 
  Automatic Semicolon Insertion & 33.33\% & 0.13 \\%& commonly used \\ 
  Arithmetic as Logic & 23.61\% & 0.02 \\%& rarely used \\ 
  Comma Operator & 15.28\% & 0.03 \\%& rarely used \\
   \bottomrule
\end{tabular}
\label{tab:occurrences-summary}
\end{table}

Considering all JavaScript projects in our dataset, we found a total of \num{364873} atom candidates, though four atoms are responsible
for 92.97\% of this total: Implicit Predicate, Post Increment, Ternary Operator, and Omitted Curly Braces. The first two, based on the results of Section~\ref{sec:survey-resuts}, can be considered atoms of confusion. The remaining atom
candidates comprise \num{25621} occurrences combined, 7.03\% of the total number of occurrences.
The Comma Operator and Arithmetic as Logic atoms are the ones that arise less
frequently (0.001\% in total with 232 and 165 occurrences, respectively).

% Regarding frequency, we proceeded with a classification of \emph{atom occurrence}
% based on thresholds used in a previous
% study~\cite{DBLP:journals/ese/MedeirosLAAKRG19}. 
% Atoms that occur 1000 times are classified as \textit{rarely used} while atoms with more than \num{10000} occurrences are classified as \textit{highly used}; \textit{commonly used} atoms occur in between.
We calculated the frequency of each atom per KLOC.
The four atoms mentioned in the previous paragraph,
 Implicit Predicate, Post Increment, Ternary Operator, and Omitted Curly Braces, occur frequently in the analyzed projects. All of them have more than 5 occurrences per KLOC on average. In particular, two atom candidates occur very frequently, Implicit Predicate (19.89 occurrences per KLOC) and Ternary Operator (10.16 occurrences per KLOC). On the other hand, Arithmetic as Logic and Comma Operator occur less than once for every 50 KLOC.
